Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
wording

We assume that the domain of the observer model is discretized, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have matrices 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, 
where $*$ represents the appropriate dimension for each matrix.
New problems arise due to the FWL truncation of these matrices. 
These are reflected in the canonical form of the resulting model (see $\mat{A}_e$ and $\mat{B}_e$),
which is expressed by:
%
{\setlength{\belowdisplayskip}{2pt}
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right], %\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
}
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_c=&\left[\begin{array}{cc}
\begin{array}{ccc}k_{p-1}-a_{p-1}&\cdots&k_1-a_1\end{array}&k_0-a_0\\
\mat{I}&\begin{array}{c}0\\ \vdots \\ 0\end{array}
\end{array}\right]&
\mat{A}_o=&\left[\begin{array}{cc}
\begin{array}{c}l_{p-1}-a_{p-1}\\ \vdots\\ l_1-a_1\end{array}&\mat{I}\\
l_0-a_0&\begin{array}{ccc}0&\cdots&0\end{array}
\end{array}\right], 
\end{align*}
}
%
where $\vec{x}'$ is the state of the model in canonical form and
$\vec{x}_e'$ is the error of the observer evaluated in the transformed
domain; further, $\mat{A}_{c}$ the closed-loop matrix of the controller in
canonical form, $\mat{A}_{o}$ the closed-loop matrix of the observer in
canonical form, $\mat{B}_{o}$ the a feedback gain matrix, $\mat{B}'$ the
noise sensitivity matrix, and $\mat{A}_{e}$ and $\mat{B}_{e}$ are the error
caused by the FWL truncation, seen after the transformation to canonical
form.

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error in the
expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in \{1, \ldots,n\},
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are
the true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop.  We note that in the case
of stability we only care that their norms are smaller than one, which is
equivalent to upper bounding the estimation of the largest eigenvalue. 
Hence, given a desired upper value $\overline{\lambda}$ (see Section
\ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to look at the columns shared with $\mat{A}_c$.
Thus we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

In many cases the response time of the system, i.e.  how fast it converges
to the set point, is a requirement of great importance.  This is given as a
``settling time''.

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p],
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
%
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }
  \overline{\lambda} = \max(|\lambda_i|) \text{ and }
  \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
%
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  For the proof see
Appendix~\ref{sec:proofs}
%
\end{theorem}

%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
%
We now use this limit to express a constraint in terms of the
coefficients of the characteristic polynomial $P_K'(z)P_L'(z)$.
Let this polynomial be:
%
\begin{equation}
P(z)=\prod_{i=0}^p (z-\lambda_i)=\sum_{j=0}^p c_jz^j.
\end{equation}
%
%We can use Jury's criterion to limit $|\lambda_i| \leq 1$.
Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
to $P'(z)=\prod_{i=0}^p (z-\lambda_i')=\sum_{j=0}^p c_j'z^j$ where $c_j'$ are
unknown.  By replacing $\lambda_i$, we obtain
%
\begin{equation}
P'(z)=\prod_{i=0}^p (z-\lambda_i')=\prod_{i=0}^p \left(z-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^p \frac{c_j}{\overline{\lambda}^{p-j}}z^j, 
\end{equation}
%
where all values are known. Thus, we can ensure that the settling time is smaller than
a certain magnitude by ensuring that 
the eigenvalues are smaller too (which we can do with Jury's criterion).

We will use $\phi_{stability}$ to refer to the combined stability and transient response specifications.

@@

In this section, we describe our technique for synthesizing safe digital
feedback controllers using CEGIS.  Broadly speaking, CEGIS denotes an
iterative technique where each step performs inductive generalization based
on counterexamples provided by a verification engine.  In each iteration, an
inductive synthesizer produces a candidate solution that obeys the
specification for a limited number of inputs.  Subsequently, this candidate
is passed to a verification phase that checks it against the specification
for all possible inputs.  If the verification fails, then the witness
counterexample is passed back to the synthesis phase, where it is added to
the set of inputs to be considered in the next loop iteration.  Otherwise,
the process terminates.

%
We enhance the standard CEGIS architecture with an
abstraction-refinement approach and with optimization, resulting in
the novel CEGIS-OR (CEGIS with Optimization and Refinement).  By using
abstraction refinement, CEGIS-OR tackles the scalability issues
inherent in program synthesis as it allows us to handle complex
systems, which we coarsely abstract and only refine on demand --
refinements are triggered by counterexamples returned by the
verification phase.
%potentially coarse abstractions of the given model that are only
%refined on demand.
Since the use of coarse abstractions may result in
low-quality candidate solutions (i.e. candidates that fail the
specification on a very large number of inputs), after each synthesis
phase we use optimization techniques to improve the produced
candidates.  In particular, our optimization procedure aims at
reducing the size of the reach space of the closed loop system, while
maintaining other dynamical properties (safety, stability, and
response time).

@@

The benchmark suite employed to evaluate the proposed technique is composed
by seven models extracted from the control literature~\cite{acrobot, cstr,
CHEN1979389, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev,
converters}.  In order to obtain statistically relevant results, we have
taken these models and modified each possible combination of their
parameters by a 10\% tolerance, thus creating seven hundred benchmarks from
this base.

@@

CEGIS-OR shows promise in exploiting modularity for improved response. 
It~however relies on the use of application-specific optimization, which
requires adequate tools for exploring different types of properties.  Our
results show significant improvement in terms of scalability over standard
CEGIS approaches, with an ability to explore complex properties, hence many
synthesis tasks may be tackled with it.  In particular we note that our
synthesis loop is more efficient for models where the solution space is
formed by connected spaces where optimization algorithms or localized
synthesis are likely to be more efficiently than global synthesis.  Finally
our application to control synthesis provides a solid proof of concept for
scalable optimal control synthesis of LTI systems.

>>>>

We assume that the domain of the observer model is discretized, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have matrices 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, 
where $*$ represents the appropriate dimension for each matrix.
New problems arise due to the FWL truncation of these matrices. 
These are reflected in the canonical form of the resulting model (see $\mat{A}_e$ and $\mat{B}_e$),
which is expressed by:
%
{\setlength{\belowdisplayskip}{2pt}
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right], %\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
}
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_c=&\left[\begin{array}{cc}
\begin{array}{ccc}k_{p-1}-a_{p-1}&\cdots&k_1-a_1\end{array}&k_0-a_0\\
\mat{I}&\begin{array}{c}0\\ \vdots \\ 0\end{array}
\end{array}\right]&
\mat{A}_o=&\left[\begin{array}{cc}
\begin{array}{c}l_{p-1}-a_{p-1}\\ \vdots\\ l_1-a_1\end{array}&\mat{I}\\
l_0-a_0&\begin{array}{ccc}0&\cdots&0\end{array}
\end{array}\right], 
\end{align*}}
%
where $\vec{x}'$ is the state of the model in canonical form and
$\vec{x}_e'$ is the error of the observer evaluated in the transformed
domain; further, $\mat{A}_{c}$ the closed-loop matrix of the controller in
canonical form, $\mat{A}_{o}$ the closed-loop matrix of the observer in
canonical form, $\mat{B}_{o}$ the a feedback gain matrix, $\mat{B}'$ the
noise sensitivity matrix, and $\mat{A}_{e}$ and $\mat{B}_{e}$ are the error
caused by the FWL truncation, seen after the transformation to canonical
form.

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error in the
expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(i.e., the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in \{1, \ldots,n\}.\,
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are
the true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop.  We note that in the case
of stability we only care that their norms are smaller than one, which is
equivalent to upper bounding the estimation of the largest eigenvalue. 
Hence, given a desired upper value $\overline{\lambda}$ (see Section
\ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to
look at the columns shared with $\mat{A}_c$.  Thus we obtain our bound
$\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$ (i.e.,
the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\
\mat{B}_e\end{array}\right]$) and apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

In many cases the response time of the system, i.e., how fast it converges
to the set point, is a high-priority requirement.  This is given as a
``settling time''.

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in \{1, \ldots,p\},
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
%
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }
  \overline{\lambda} = \max(|\lambda_i|) \text{ and }
  \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
%
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  The proof is in
Appendix~\ref{sec:proofs}.
%
\end{theorem}

%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
%
We now use this limit to express a constraint in terms of the
coefficients of the characteristic polynomial $P_K'(z)P_L'(z)$.
Let this polynomial be:
%
\begin{equation}
P(z)=\prod_{i=0}^p (z-\lambda_i)=\sum_{j=0}^p c_jz^j.
\end{equation}
%
%We can use Jury's criterion to limit $|\lambda_i| \leq 1$.
Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$, which can be done by applying Jury's criterion
to $P'(z)=\prod_{i=0}^p (z-\lambda_i')=\sum_{j=0}^p c_j'z^j$ where $c_j'$ are
unknown.  By replacing $\lambda_i$, we obtain
%
\begin{equation}
P'(z)=\prod_{i=0}^p (z-\lambda_i')=\prod_{i=0}^p \left(z-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^p \frac{c_j}{\overline{\lambda}^{p-j}}z^j, 
\end{equation}
%
where all values are known.  Thus, we can ensure that the settling time is
smaller than a certain magnitude by ensuring that the eigenvalues are
smaller too (which we can do with Jury's criterion).

We will use $\phi_\mathit{stability}$ to refer to the combined stability and
transient response specifications.

@@

In this section, we describe our technique for synthesizing safe digital
feedback controllers.  We use CEGIS, which is an iterative technique where
each step performs inductive generalization based on counterexamples
provided by a verification engine.  In each iteration, an inductive
synthesizer produces a candidate solution that obeys the specification for a
limited number of inputs.  Subsequently, this candidate is passed to a
verification phase that checks it against the specification for all possible
inputs.  If the verification fails, then the witness counterexample is
passed back to the synthesis phase, where it is added to the set of inputs
to be considered in the next loop iteration.  Otherwise, the process
terminates.

We enhance the standard CEGIS architecture with an abstraction-refinement
approach and with optimization.  We call the resulting new technique
\emph{CEGIS-OR} (CEGIS with Optimization and Refinement).  By using
abstraction refinement, CEGIS-OR addresses the scalability issues inherent
in program synthesis and enables us to handle complex systems, which we
coarsely abstract and only refine on demand -- refinements are triggered by
counterexamples returned by the verification phase.
%
%potentially coarse abstractions of the given model that are only
%refined on demand.
%
Since the use of coarse abstractions may result in low-quality candidate
solutions (i.e.  candidates that fail the specification on a very large
number of inputs), after each synthesis phase we use optimization techniques
to improve the produced candidates.  In particular, our optimization
procedure aims at reducing the size of the reach space of the closed loop
system, while maintaining other dynamical properties (safety, stability, and
response time).

@@

The benchmark suite employed to evaluate the proposed technique is composed
by seven models extracted from the control literature~\cite{acrobot, cstr,
CHEN1979389, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev,
converters}.  In~order to obtain statistically relevant results, we have
taken these models and modified each possible combination of their
parameters by a 10\% tolerance, thus creating seven hundred benchmarks from
this base.

@@

CEGIS-OR shows promise in exploiting modularity for improved response. 
It~however relies on the use of application-specific optimization, which
requires adequate tools for exploring different types of properties.  Our
results show significant improvements in terms of scalability over standard
CEGIS approaches, with an ability to explore complex properties, hence many
synthesis tasks may be solved with it.  In particular, we note that our
synthesis loop is more efficient for models where the solution space is
formed by connected spaces where optimization algorithms or localized
synthesis are likely to be more efficiently than global synthesis.  Finally,
our application to control synthesis provides a solid proof of concept for
scalable optimal control synthesis of LTI systems.
